Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-03-30 Writing/manuscript_cea.rmd
additional changes

Figure \@ref(fig:vis-MB-BRF-y) illustrates the distinction between yield prediction at observed points and yield response function prediction, giving insights into why good yield prediction does not necessarily lead to good EONR prediction in general. The black curves are the true yield response functions for two sites in a single simulation round. Red symbols represent true (circle) or predicted (triangle) yields at the rate observed in the data: the rate that was actually applied in the experiment. That is, 150 kg/ha was applied to the two sites in the experiment. Blue symbols represent true (circle) or predicted (triangle) yield at some rates that were not observed (or applied) for the two sites in the experiment. Prediction-oriented ML methods train models to minimize errors in yield level prediction at the "observed" rates (red points). Both RF- and BRF-trained models predict yield level at the observed rate well for the two sites (the vertical distance between the red circle and triangle). However, as you move away from the observed N rate, yield prediction errors tend to become larger (of course smaller yield prediction errors can be observed at unobserved rates some times). More important, connecting the predicted yield points does not trace the true yield response curves well for both RF and BRF. Specifically, in this example, the predicted impact of N on yield from RF and BRF are both smaller than the true impact of N.

@@

In essence, Figure \@ref(fig:vis-MB-BRF-y) illustrates that good performance in predicting yield "levels" at the "observed" points does not mean good performance in predicting yield "response" function, which is about finding out what "would" happen to yield if rates other than the one actually applied were applied. This happens because you can predict yield "levels" for wrong reasons by giving wrong amount of credit to wrong drivers (variables) of yield levels. The training process of prediction-oriented ML methods may twist the reasonings of why yield varies in a way that sacrifices its ability to recover yield response function as long as it fits its primary objective of predicting yield levels well at the "observed" nitrogen rate for each site well. This problem is also observable at Figure \@ref(fig:plot-tre), which shows both RF and BRF underestimates the impact of N on yield: or in other words, their predicted site-specific yield response functions tend to be flatter than they truly are. Of course, the CNN approach in our study is an extreme case where it completely fails to recognize the contribution of N on yield. But, remember that CNN does performs almost as good as RF and BRF in yield prediction as shown in Table \@ref(tab:table-y).

>>>>

Figure \@ref(fig:vis-MB-BRF-y) illustrates the distinction between yield prediction at observed points and yield response function prediction, giving insights into why good yield level prediction does not necessarily lead to good EONR prediction in general. The black curves are the true yield response functions for two sites in a single simulation round. Red symbols represent true (circle) or predicted (triangle) yields at the rate observed in the data: the rate that was actually applied in the experiment. That is, 150 kg/ha was applied to the two sites in the experiment. Blue symbols represent true (circle) or predicted (triangle) yield at some rates that were not observed (or applied) for the two sites in the experiment. Prediction-oriented ML methods train models to minimize errors in yield level prediction at the "observed" rates (red points). Both RF- and BRF-trained models predict yield level at the observed rate well for the two sites (the vertical distance between the red circle and triangle). However, as you move away from the observed N rate, yield prediction errors tend to become larger (of course smaller yield prediction errors can be observed at unobserved rates some times). More important, connecting the predicted yield points does not trace the true yield response curves well for both RF and BRF. Specifically, in this example, the predicted impact of N on yield from RF and BRF are both smaller than the true impact of N.

@@

In essence, Figure \@ref(fig:vis-MB-BRF-y) illustrates that good performance in predicting yield "levels" at the "observed" points does not mean good performance in predicting yield "response" function, which is about finding out what "would" have happened to yield if rates other than the one actually applied were applied. This happens because you can predict yield "levels" for wrong reasons by giving wrong amount of credit to wrong drivers (variables) of yield levels. The training process of prediction-oriented ML methods may twist the reasonings of why yield varies in a way that sacrifices its ability to recover yield response function as long as it fits its primary objective of predicting yield levels well at the "observed" nitrogen rate for each site well. This problem is also observable at Figure \@ref(fig:plot-tre), which shows both RF and BRF underestimates the impact of N on yield: or in other words, their predicted site-specific yield response functions tend to be flatter than they truly are. Of course, the CNN approach in our study is an extreme case where it completely fails to recognize the contribution of N on yield. But, remember that CNN does performs almost as good as and sometimes surpass RF and BRF in yield prediction as shown in Table \@ref(tab:table-y).
